---
title: "Marketing Analytics Final Project"
author: "Harsh Tandon, Tianqing Feng, Bingbing Pan"
output:
  word_document: default
  pdf_document: default
  html_document: default
subtitle: Bestbuy Case Study
---
- Load file and packages 
```{r}
#setwd('/Users/bing/Desktop/COURSEWORK/MarketingAnalytics_2020/DataSets for student projects')
library(data.table)
library(MASS)
library(ggplot2)
library(caret)
library(e1071)
library(stargazer)
library(caTools)
library(randomForest)

df <- read.csv('BestBuy.csv')
stargazer(df, 
          title = "Statistical Summary", type = "text",
          digits = 3)
```
- Explore data 
```{r}
cormat <- round(cor(df),2)
head(cormat)
library(reshape2)
melted_cormat <- melt(cormat)
head(melted_cormat)
library(ggplot2)

# Get lower triangle of the correlation matrix
get_lower_tri<-function(cormat){
  cormat[upper.tri(cormat)] <- NA
  return(cormat)
}
# Get upper triangle of the correlation matrix
get_upper_tri <- function(cormat){
  cormat[lower.tri(cormat)]<- NA
  return(cormat)
}

# Reorder the correlation matrix
upper_tri <- get_upper_tri(cormat)

# Melt the correlation matrix
library(reshape2)
melted_cormat <- melt(upper_tri, na.rm = TRUE)
# Heatmap
library(ggplot2)


reorder_cormat <- function(cormat){
  # Use correlation between variables as distance
  dd <- as.dist((1-cormat)/2)
  hc <- hclust(dd)
  cormat <-cormat[hc$order, hc$order]
}

cormat <- reorder_cormat(cormat)
# Melt the correlation matrix
melted_cormat <- melt(upper_tri, na.rm = TRUE)
# Create a ggheatmap
ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Pearson\nCorrelation") +
  theme_minimal()+ # minimal theme
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1))+
  coord_fixed()

# print ggheatmap
ggheatmap + 
  geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    legend.justification = c(1, 0),
    legend.position = c(0.6, 0.7),
    legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                               title.position = "top", title.hjust = 0.5))


```
Prepare data for later analysis
```{r}
# Change dummy variables to factors
df$married <- as.factor(df$married)
df$MyBestBuy <- as.factor(df$MyBestBuy)
df$appliances <- as.factor(df$appliances)
df$newcustomer <- as.factor(df$newcustomer)
df$weekend <- as.factor(df$weekend)
df$Warranty <- as.factor(df$Warranty)
df$hisp <- as.factor(df$hisp)
df$hhincome <- log(1+df$hhincome)
str(df)
stargazer(df, 
          title = "Statistical Summary", type = "text",
          digits = 3)
```
- Split data into Training and Testing
```{r}
split = sample.split(df$Warranty, SplitRatio = 0.8)
#training_set <- 1:2560
#test_set <- !(1:nrow(df) %in% training_set) 
training_set = subset(df, split == TRUE)
test_set = subset(df, split == FALSE)
```
- Train Random Forest Model
```{r}
classifier = randomForest(x = training_set[-9],
                          y = training_set$Warranty,
                          ntree = 500)
```
- Printing more detailed prediction accuracy and decision tree
```{r}
library(party)
y_pred = predict(classifier, newdata = test_set[-c(9)], type = "response")
confusionMatrix(y_pred, test_set$Warranty)
tree <- ctree(Warranty ~., data = test_set, controls = ctree_control(mincriterion = 0.30))
tree
plot(tree)
```
- Print AUC score for random forest model
```{r}
library(pROC)
is.numeric(y_pred)
y_pred2 = as.numeric(y_pred)
is.numeric(y_pred2)
auc(test_set$Warranty, y_pred2)

```
- Train Logistic Models
```{r warning=FALSE}
# The first model (abandon productgeneration (highly correlated with PriceCategory), personid)
glm1 <- glm(Warranty ~ age + hisp + PriceCategory + married + MyBestBuy + hhincome + appliances + familysize + newcustomer + weekend, data = training_set, family="binomial")
stargazer(glm1, title = "Logit Result", type = "text", column.labels = c("Model-1"), df = FALSE, digits = 3, star.cutoffs = c(0.05, 0.01, 0.001))

#The second model (remove newcustomer and weekend)
glm2 <- glm(Warranty ~ age + hisp + PriceCategory + married + MyBestBuy + hhincome + appliances + familysize, data = training_set, family="binomial")
stargazer(glm1, glm2, title = "Logit Result", type = "text", column.labels = c("Model-1", "Model-2"), df = FALSE, digits = 3, star.cutoffs = c(0.05, 0.01, 0.001))

#The third model (remove  familysize)
glm3 <- glm(Warranty ~ age + hisp + PriceCategory + married + MyBestBuy + hhincome + appliances, data = training_set, family="binomial")
stargazer(glm1, glm2, glm3, title = "Logit Result", type = "text", column.labels = c("Model-1", "Model-2", "Model-3"), df = FALSE, digits = 3, star.cutoffs = c(0.05, 0.01, 0.001))

#The fourth model - interaction term (newcustomer * MyBestBuy)
glm4 <- glm(Warranty ~ age + newcustomer * MyBestBuy + hisp + PriceCategory + appliances + married + hhincome, data = training_set, family="binomial")
stargazer(glm1, glm2, glm3, glm4, title = "Logit Result", type = "text", column.labels = c("Model-1", "Model-2", "Model-3", "Model-4"), df = FALSE, digits = 3, star.cutoffs = c(0.05, 0.01, 0.001))


#The fifth model - interaction term (PriceCategory * appliances) - keep
glm5 <- glm(Warranty ~ age + hisp + PriceCategory * appliances + married + MyBestBuy + hhincome, data = training_set, family="binomial")
stargazer(glm1, glm2, glm3, glm4, glm5, title = "Logit Result", type = "text", column.labels = c("Model-1", "Model-2", "Model-3", "Model-4", "Model-5"), df = FALSE, digits = 3, star.cutoffs = c(0.05, 0.01, 0.001))

#The sixth model - two interaction terms (newcustomer * mybestby + PriceCategory * appliances)
glm6 <- glm(Warranty ~ newcustomer * MyBestBuy + hisp + PriceCategory * appliances + married + hhincome, data = training_set, family="binomial")
stargazer(glm1, glm2, glm3, glm4, glm5, glm6, title = "Logit Result", type = "text", column.labels = c("Model-1", "Model-2", "Model-3", "Model-4", "Model-5", "Model-6"), df = FALSE, digits = 3, star.cutoffs = c(0.05, 0.01, 0.001))
```
- Predict Test_set results using model 5
```{r}
predict_y <- predict(glm5, newdata = test_set[-9], type="response")
thresh  <- 0.5            
predFac <- cut(predict_y, breaks=c(-Inf, thresh, Inf),labels=c("0","1"))
confusionMatrix(predFac, test_set$Warranty)
```
- Print AUC score for logist model
```{r}
is.numeric(predFac)
pred_fac = as.numeric(predFac)
is.numeric(pred_fac)
auc(test_set$Warranty, pred_fac)
```
- Graphs of variables
```{r}
sub_mar=subset(df,select = c('married','Warranty'))
h1=as.data.frame(table(sub_mar))
plot1=ggplot(h1,aes(x=married,y=Freq,fill=Warranty)) +
  geom_bar(position = "dodge", stat = "identity")+
  geom_text(aes(label = Freq),vjust = 1.5, colour = "white", position = position_dodge(.9), size = 5)


sub_MBB=subset(df,select = c('MyBestBuy','Warranty'))
h2=as.data.frame(table(sub_MBB))
plot2=ggplot(h2,aes(x=MyBestBuy,y=Freq,fill=Warranty)) +
  geom_bar(position = "dodge", stat = "identity")+
  geom_text(aes(label = Freq),vjust = 1.5, colour = "white", position = position_dodge(.9), size = 5)

sub_nc=subset(df,select = c('newcustomer','Warranty'))
h3=as.data.frame(table(sub_nc))
plot3=ggplot(h3,aes(x=newcustomer,y=Freq,fill=Warranty)) +
  geom_bar(position = "dodge", stat = "identity")+
  geom_text(aes(label = Freq),vjust = 1.5, colour = "white", position = position_dodge(.9), size = 5)

sub_his=subset(df,select = c('hisp','Warranty'))
h4=as.data.frame(table(sub_his))
plot4=ggplot(h4,aes(x=hisp,y=Freq,fill=Warranty)) +
  geom_bar(position = "dodge", stat = "identity")+
  geom_text(aes(label = Freq),vjust = 1.5, colour = "white", position = position_dodge(.9), size = 5)

plot5=ggplot(df, aes(x = hhincome)) +
  geom_histogram(binwidth = 0.5, fill = "lightblue", colour = "black")

library(gridExtra)

comb_graphs<-grid.arrange(plot1,plot2,plot3,plot4,plot5, ncol = 2)
comb_graphs
```
- Graph for interaction term
```{r}
## interaction term visualization
#setwd('/Users/bing/Desktop/COURSEWORK/MarketingAnalytics_2020/DataSets for student projects')
#install.packages('ggeffects')
library(ggeffects)
df <- read.csv('BestBuy.csv')
df$married <- as.factor(df$married)
df$MyBestBuy <- as.factor(df$MyBestBuy)
df$appliances <- as.factor(df$appliances)
df$newcustomer <- as.factor(df$newcustomer)
df$weekend <- as.factor(df$weekend)
df$hisp <- as.factor(df$hisp)
df$hhincome <- log(1+df$hhincome)
linear<- lm(Warranty ~ appliances*PriceCategory, data=df)

meffects <- ggpredict(linear, terms=c("PriceCategory", "appliances")) # generates a tidy data frame  

ggplot(meffects,aes(x, predicted, colour=group)) + geom_line(size=1.3) + 
    xlab("PriceCategory") + ylab("Warranty") +
    labs(colour="Appliance?") + 
    scale_colour_discrete(labels=c("No", "Yes"))

```

